---
title: "Animal_outcomes"
author: "Aura Zuluaga"
date: "12/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(janitor)
library(skimr)
library(ggplot2)
library(knitr)
library(dplyr)
theme_light()
```

Structure of the report:
1.	Executive summary
2.	Introduction/Background (maximum 100 words)
3.	Research objective (main objective &/or sub-objectives) & questions.

## Possible Objectives:

### Reduce the euthanise levels per state territory
### Increase the percentage of adoption in animals per state

4.	Review information

5.	Analysis and evaluate

```{r}
animal_outcomes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-21/animal_outcomes.csv')
clean_names(animal_outcomes)
animal_outcomes$animal_type<-as.factor(animal_outcomes$animal_type)
animal_outcomes$outcome<-as.factor(animal_outcomes$outcome)
animal_outcomes$year<-as.numeric(animal_outcomes$year)
animal_outcomes[is.na(animal_outcomes)] = 0
skim(animal_outcomes)
```

# 1. How many animals have been released/rehomed/euthanised/currently_in_care from 1999 to 2018 in Australia? How the outcomes' percentage levels have changed in dogs,cats and wildlife ? 

###COMPARISON ANIMAL OUTCOMES 

```{r}
table1<-animal_outcomes%>%count(year,outcome, sort=TRUE, wt=Total)%>%
    mutate(outcome=fct_lump(outcome, 7),
           outcome=fct_reorder(outcome, -n, sum)) 
           kable(table1)
table1%>%ggplot(aes(year, n,color=outcome)) +
    geom_line() +
labs(
    x = "Year",
    y = "Number of Animals",
    title="Comparison of outcomes of all animals through time",
    fill = NULL)+
    facet_wrap(~outcome,scales="free_y")
```

###COMPARISON OUTCOMES DOGS

```{r}
dogs1<-animal_outcomes%>%
filter(animal_type=="Dogs")%>%
count(year,animal_type, outcome, wt=Total)%>%
mutate(outcome=fct_lump(outcome, 4),
           outcome=fct_reorder(outcome, n, sum)) 
kable(dogs1)

dogs1%>%ggplot(aes(year, n, fill= outcome)) +
  geom_bar(position = "fill", stat = "identity",colour = "black")+
scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(palette="Pastel1")+
labs(
    x = "Year",
    y = "Percentage of Dogs",
    title="Comparison of outcomes of dogs through time",
    fill = NULL)
```


###COMPARISON OUTCOMES CATS

```{r}
cats1<-animal_outcomes%>%
filter(animal_type=="Cats")%>%
count(year,animal_type, outcome, wt=Total)%>%
mutate(outcome=fct_lump(outcome, 4),
           outcome=fct_reorder(outcome, n, sum)) 
kable(cats1)

cats1%>%ggplot(aes(year, n, fill= outcome)) +
  geom_bar(position = "fill", stat = "identity",colour = "black")+
scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(palette="Pastel1")+
labs(
    x = "Year",
    y = "Percentage of Cats",
    title="Comparison of outcomes of cats through time",
    fill = NULL)
```


###COMPARISON OUTCOMES WILDLIFE

```{r}
wild1<-animal_outcomes%>%
filter(animal_type=="Wildlife")%>%
count(year,animal_type, outcome, wt=Total)%>%
mutate(outcome=fct_lump(outcome, 4),
           outcome=fct_reorder(outcome, n, sum)) 
kable(wild1)

wild1%>%ggplot(aes(year, n, fill= outcome)) +
  geom_bar(position = "fill", stat = "identity",colour = "black")+
scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(palette="Pastel1")+
labs(
    x = "Year",
    y = "Percentage of Wildlife",
    title="Comparison of outcomes of wildlife through time",
    fill = NULL)
```

# 2. Identify how the number of euthanized wildlife animals has changed during the last decade? Does it have a significant change through years?

### WILDLIFE EUTHANIZED

```{r}
table2<-animal_outcomes%>%filter(animal_type=='Wildlife',outcome=='Euthanized')%>%
count(year,animal_type,outcome,wt=Total)
kable(table2)


table2%>%ggplot(aes(year, n)) +
  geom_line(colour="purple")+
labs(
    x = "Year",
    y = "Number of Wildlife Euthanized",
    title="Number of Wildlife Euthanized through time",
    fill = NULL)
```

##PERCENTAGE OF WILDLIFE EUTHANIZED IN AUSTRALIA

```{r}
table2b<-animal_outcomes%>%filter(animal_type=='Wildlife')%>%
count(year,animal_type,wt=Total,name="nt")
kable(table2b)


merge2 <- merge(table2,table2b,by=c("year","animal_type"))%>%
mutate(table2b, p = (n/nt))
kable(merge2)


merge2%>%ggplot(aes(year, p)) +
    geom_line(color="red") +
scale_y_continuous(labels = scales::percent_format()) +
labs(
    x = "Year",
    y = "% of Wildlife Euthanized",
    title="Percentage of Wildlife Euthanized in Australia through time",
    fill = NULL)
```


# 3. What regions of Australia euthanised more animals?

```{r}
euth3<-animal_outcomes%>%
filter(outcome=="Euthanized")%>%
pivot_longer(ACT:WA, names_to ="state", values_to="n")%>%
group_by(year,state)%>%
summarise(n =sum(n))
kable(euth3)


euth3%>%ggplot(aes(year, n,color=state)) +
    geom_line() +
labs(
    x = "Year",
    y = "Number of Euthanized Animals",
    title="Comparison of numbers of euthanized animals by state through time",
    fill = NULL)
```


# 4. What are the adoption levels per state territory?

##NUMBER OF ANIMALS ADOPTED OR RECLAIMED BY STATE

```{r}
adopt4<-animal_outcomes%>%
filter(outcome != "Other", outcome != "Euthanized", outcome != "Currently In Care", outcome != "Transferred", outcome != "Released", outcome != "In Stock")%>%
pivot_longer(ACT:WA, names_to ="state", values_to="n")%>%
group_by(year,state)%>%
summarise(n =sum(n))
kable(adopt4)


adopt4%>%ggplot(aes(year, n,color=state)) +
    geom_line() +
labs(
    x = "Year",
    y = "Number of Adopted Animals",
    title="Comparison of numbers of adopted animals by state through time",
    fill = NULL)
```

##PERCENTAJE OF ANIMALS ADOPTED OR RECLAIMED BY STATE

```{r}
animals4<-animal_outcomes%>%
pivot_longer(ACT:WA, names_to ="state", values_to="n")%>%
group_by(year,state)%>%
summarise(nt =sum(n))
kable(animals4)


merge4 <- merge(adopt4,animals4,by=c("year","state"))%>%
mutate(animals4, p = (n/nt))
kable(merge4)


merge4%>%ggplot(aes(year, p, color= state)) +
    geom_line() +
scale_y_continuous(labels = scales::percent_format()) +
labs(
    x = "Year",
    y = "% of animales adopted or reclaimed",
    title="Percentage of animales adopted or reclaimed by state through time",
    fill = NULL)+
    facet_wrap(~state,scales="free_y")
```

##TOTAL OF ANIMALS THAT GO INTO DE SYSTEM PER STATE

```{r}
animals4%>%ggplot(aes(year, nt,color=state)) +
    geom_line() +
labs(
    x = "Year",
    y = "Number of Animals",
    title="Total of animals that go into the system per state",
    fill = NULL)
```

##ANIMALS ADOPTED IN THE COUNTRY

```{r}
country4<-animal_outcomes%>%
filter(outcome != "Other", outcome != "Euthanized", outcome != "Currently In Care", outcome != "Transferred", outcome != "Released", outcome != "In Stock")%>%
pivot_longer(ACT:WA, names_to ="state", values_to="n")%>%
group_by(year)%>%
summarise(n =sum(n))
kable(country4)


country4%>%ggplot(aes(year, n)) +
    geom_line() +
labs(
    x = "Year",
    y = "Number of Adopted Animals",
    title="Comparison of numbers of adopted animals in the country through time",
    fill = NULL)
```

## PERCENTAGE OF ANIMALS ADOPTED IN THE COUNTRY

```{r}
country4b<-animal_outcomes%>%
pivot_longer(ACT:WA, names_to ="state", values_to="n")%>%
group_by(year)%>%
summarise(nt =sum(n))
kable(country4b)


merge4b <- merge(country4,country4b,by=c("year"))%>%
mutate(country4, p = (n/nt))
kable(merge4b)


merge4b%>%ggplot(aes(year, p)) +
    geom_line(colour="blue") +
scale_y_continuous(labels = scales::percent_format()) +
labs(
    x = "Year",
    y = "% of animales adopted or reclaimed",
    title="Percentage of animales adopted or reclaimed in Australia through time",
    fill = NULL)
```


# 5. WhaT type of animal is more likely to be adopted in Australia?

```{r}
table5<-animal_outcomes%>%filter(outcome=='Rehomed')%>%
count(year,animal_type,outcome,wt=Total)%>%
mutate(animal_type=fct_reorder(animal_type, n, sum)) 
kable(table5)


table5%>%ggplot(aes(year, n, fill= animal_type)) +
  geom_bar(position = "fill", stat = "identity",colour = "black")+
scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(palette="Pastel2")+
labs(
    x = "Year",
    y = "Percentage of Animals Adopted",
    title="Comparison of the percentage of animals adopted in Australia through time",
    fill = NULL)

```


6.	Findings and discussion

# Predict # of adoptions in the coming years. What are the more critical 
# 5. Why are the levels of adoption lower in NSW than in other Australian states?

7.	Recommendations
8.	Limitations of this study


9.	Conclusion
10.	References
